---
title: "Data Mining - Assignment 1"
date: "October 7, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Assignment
A restaurant owner has a core menu that has remained constant throughout the past two years, while many
other dishes have rotated on and off the menu. The owner has prepared a dataset of all orders containing
elements of that core menu in which the customer also ordered wine. The dataset has 3 columns: orderNumber, order, and type.

orderNumber identifies each individual person's order at a table, order itemizes the order's components, and type provides the type of item in the order (takes one of 3 values: 'Meat', 'Side', 'Wine').

The restaurant owner would like to add wine suggestions to the menu next to each meat and analyze her
customer ordering habits in general. She is also interested in showing a picture of a popular meal of hers in an ad in Gourmet Diner magazine. Your job is to address her interests and also to report anything else of interest in terms of customer ordering habits, all in a nice summary report that is complete but to the point.


```{r, include=F}
###loading Packages
library(tidyverse)
library(haven)
library(MASS)
library(brglm)
library(arm)
library(multcomp)
library(lsmeans)
library(arules)
library(xtable)
```

###Load Data
```{r}
dir = "C:\\Users\\sbessias\\Google Drive\\NCSU\\Data Mining\\HW Data\\"
rest = read_csv(paste(dir, "restaurantData.csv", sep = ""))
```

###Check for missing values and duplicates
There do not appear to be any missing values.
There are 388 orderNumbers that appear more than 3 times, but these aren't necessarily duplicates. Some people may have ordered multiple meals. Leaving them in the dataset for now.
```{r}
table(rest$order, useNA='ifany')
table(rest$order[rest$type=='Side'])
table(rest$order[rest$type=='Meat'])
table(rest$order[rest$type=='Wine'])
summary(rest$orderNumber, useNA='ifany')
s <- table(rest$orderNumber)
#sort(s[s>3])
dim(s[s>3])
```

###Read in Transactions
```{r, include=F}
restaurant = read.transactions(file=paste(dir, "restaurantData.csv", sep=""), 
                      format='single', 
                      cols=c("orderNumber", "order"),
                      sep = ",")

#Format should be specified as single because each line in our dataset corresponds to a single item and contains ids for the transaction and the item.
summary(restaurant)
```

###Examine transactions and visualize item frequencies
```{r}
inspect(restaurant[1:5]) #see first five transactions

#see frequencies of all the items, in order of most frequent to least
sort(itemFrequency(restaurant), decreasing = TRUE)
freqs <- sort(itemFrequency(restaurant), decreasing = TRUE)

itemFr

#save frequencies to .csv for plotting later
write.csv(freqs, file=paste(dir, 'freqs.csv', sep=''))
itemFrequencyPlot(restaurant, topN=14, horiz=TRUE, xlab="Relative Item Frequency")

#Visualize the sparse matrix for the first 100 transactions
image(restaurant[1:100])
```

###Create Association Rules 
```{r}
rules <- apriori(restaurant, parameter = list(support = 0.006, confidence = 0.25, minlen = 1))
summary(rules)

#view top 20 association rules, sorted by lift value
allrules <- as(rules, "data.frame")
top20 <- allrules[1:20,]
write.csv(top20, file=paste(dir, "top20rules.csv"))
write.csv(allrules, file=paste(dir, "allrules.csv"))

#write.csv(x=inspect(rules), file=paste(dir, 'rulestable.csv', sep=''))
write.csv(allrules, file=paste(dir, 'rulestable.csv', sep=','))

############### make a nicer plot  ################

#load relative frequencies
freqs <- read.csv(file=paste(dir, 'freqs.csv', sep=''), sep=',', header=T)
freqs

#add row for category
Category <- c("Meat", "Wine", "Wine", "Side", "Side", "Meat", "Meat", "Side", "Wine", 
              "Side", "Side", "Side", "Meat", "Wine")
freqs <- cbind(freqs, Category)
freqs

#add headers
colnames(freqs) <- c("Menu.Item", "Relative.Frequency", "Category")


#plot results
ggplot(data=freqs, aes(x=reorder(Menu.Item, Relative.Frequency), 
                       y=Relative.Frequency, fill=Category)) +
    geom_bar(stat="identity") + 
    theme_classic() + 
    coord_flip() +
    xlab("Relative Frequency") + ylab("Menu Item") +
    ggtitle("Customer orders by relative frequency (popularity)\n")

```

###Look at rules for each of the meats to find wine parings
```{r}

#Rules with Filet Mingon in them
filetrules <- subset(rules, items %in% c("Filet Mignon"))
inspect(sort(filetrules, by = "lift"))

#Rules with Roast Chicken in them
chickenrules <- subset(rules, items %in% c("Roast Chicken"))
inspect(sort(chickenrules, by = "lift"))

#Rules with Pork Tenderloin in them
porkrules <- subset(rules, items %in% c("Pork Tenderloin"))
inspect(sort(porkrules, by = "lift"))

#Rules with Duck in them
duckrules <- subset(rules, items %in% c("Duck"))
inspect(sort(duckrules, by = "lift"))
```
